---
title: "Charlie Baker vs Martha Coakley, 2014"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: spacelab
---

```{r parameters, echo=FALSE,include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo=FALSE,fig.width=10, fig.height=8, comment=NA, warning=FALSE, message=FALSE)

library(htmltools)
library(leaflet)
library(maptools)
library(sp)
library(stringr)
library(tidyverse)
```

```{r data,cache=TRUE}
# Town shapefile from
# http://www.mass.gov/anf/research-and-tech/it-serv-and-support/application-serv/office-of-geographic-information-massgis/datalayers/townsurvey.html
towns = raster::shapefile('townssurvey_shp/TOWNSSURVEY_POLYM.shp')
towns = spTransform(towns, CRS("+init=EPSG:4326"))
towns_simp = rmapshaper::ms_simplify(towns)
# Election results from
# http://electionstats.state.ma.us/elections/view/35901/
results = read_csv('PD43+__2014_Governor_General_Election.csv') %>% 
  select(-X2, -X3) %>% 
  mutate(TOWN=str_to_upper(`City/Town`) %>% 
           str_replace('E\\.', 'EAST') %>% 
           str_replace('W\\.', 'WEST') %>% 
           str_replace('S\\.', 'SOUTH') %>% 
           str_replace('N\\.', 'NORTH') %>% 
           str_replace('MANCHESTER-BY-THE-SEA', 'MANCHESTER')) %>% 
  filter(TOWN != 'TOTALS')

d = merge(towns_simp, results, by='TOWN')

# Place dots, one per hundred votes
# Round up so towns with few voters get some dots
red = dotsInPolys(d, ceiling(d$`Baker/ Polito`/100))
blue = dotsInPolys(d, ceiling(d$`Coakley/ Kerrigan`/100))

make_popups = function(d) {
   lapply(1:nrow(d),
       function(i) {
         row = d[i,]
   HTML(as.character(
     tags$table(tags$caption(strong(row$`City/Town`)),
     tags$tr(tags$td('Baker'), 
             tags$td(row$`Baker/ Polito`),
             tags$td(scales::percent(row$`Baker/ Polito`/row$`Total Votes Cast`))),
     tags$tr(tags$td('Coakley'), 
             tags$td(row$`Coakley/ Kerrigan`),
             tags$td(scales::percent(row$`Coakley/ Kerrigan`/row$`Total Votes Cast`)))
   )))
   })
  }

d$popup = make_popups(d@data)
```

The Vote
============

Row
-----------------------------------------------------------------------

### Votes for Charlie Baker and Martha Coakley, 2014 Massachusetts Gubernatorial Election

```{r}
leaflet() %>% addPolygons(data=d, label=~`City/Town`, popup=~popup,
                          color='gray', 
                          fillColor='#F0F0F0', fillOpacity=1,
                          weight=1) %>% 
  addCircleMarkers(data=red, radius=1, stroke=FALSE, 
                   fillColor='red', fillOpacity=0.5, group='Baker') %>% 
  addCircleMarkers(data=blue, radius=1, stroke=FALSE, 
                   fillColor='blue', fillOpacity=0.5, group='Coakley') %>% 
  addLayersControl(overlayGroups=c('Baker', 'Coakley'))
```

About
================

### About this map

This map shows the number of votes received by Charlie Baker and 
Martha Coakley in the 2014 Massachusetts Gubernatorial Election.

Each dot represents 100 votes. Click on a town for additional information.

Data is from the
<a href="http://electionstats.state.ma.us/elections/view/35901/" target="_blank">Massachusetts Secretary of the Commonwealth</a>, 
mapped by Kent S Johnson. 

<span style='float:right;font-style: italic;'>`r Sys.Date()`</span></small>
  